---
title: "Coffee"
author: "Erica Galvan"
date: "2024-11-12"
output: html_document
---
*Introduction*


*Load Packages and Data*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
library(tidyverse)
library(ggplot2)
library(dplyr)
```

*Data Source*
-Link to Data
-Explaning the source and concepts of how the survey was taken 
-Explain the experiment that was done

*Questions*
Who typically drinks coffee?
-What does the consumer generally add to their coffee?
-Where do they drink coffee?
Why do they drink?
-Compare the surveyers' preferences based on the results 

```{r}
#HOW DOES THE CONSUMER DRINK COFFEE 


```

```{r}
#QUESTION: WHAT IS THE PREFERRED ROAST LEVEL?" * add counts to the amount 
coffee_survey %>% 
  filter(!is.na(prefer_abc)) %>% 
  ggplot(aes(x = prefer_abc, fill = prefer_abc)) + geom_bar() + scale_fill_manual( values = c("tan", "tan3", "saddlebrown"))+
  labs(title = "Coffee Preference: A, B, or C?", 
       x = NULL,
       y = "Count")+ 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.3, color = "black", size = 4)
```
  

```{r}
#QUESTION: WHAT IS THE PREFERENCE BETWEEN A / D? * ADD COUNT AMOUNTS 
coffee_survey %>% 
  filter(!is.na(prefer_ad)) %>% 
  ggplot(aes(x = prefer_ad, fill = prefer_ad)) + geom_bar() + scale_fill_manual( values = c("tan", "peachpuff3"))+
  labs( title = "Coffee Preference: A or D?",
    x = NULL,
    y = "Count"
  )+ geom_text(stat = "count", aes(label = ..count..), vjust = -0.3, color = "black", size = 4)


```


```{r}
#Overall Preference Between them all?

 coffee_survey%>% 
 filter(!is.na(prefer_overall)) %>%
 ggplot(aes(prefer_overall)) +
 geom_bar(fill= c("tan", "tan3", "saddlebrown", "peachpuff3")) +
 labs(y='Count',
      x = NULL,
      title='Preferred Coffee Between A, B, C or D?')+ 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.3, color = "black", size = 4)



```


```{r}
#WHAT DOES THE CONSUSMER ADD TO THE COFFEE 



```


```{r}




```


```{r}




```


```{r}




```


*NOTES*
Introduce what the data is(where it comes from)
-Why is it important?
-What features?
*Next, highlight the questions you will be asking.


*Questions*
Where do people brew their coffee? How is it brewed (make relation?)

What is the most popular coffee roast(gender + age)?

#Between Coffee A, Coffee B, and Coffee C which did you prefer?
#Between Coffee A and Coffee D, which did you prefer?
#Favorite Overall Coffee

What are the preferences between popular coffee between gender?

