---
title: "Great American Coffee Taste Test "
author: "Erica Galvan"
date: "2024-11-26"
output:
  slidy_presentation:
    fig_height: 8
    fig_width: 10
  pdf_document: default
 
---
# Introduction
* Coffee expert, "World Champion Barista" and Youtuber 'James Hoffman' partnered with Cometeer, a flash frozen coffee pod delivery service.  
* Live coffee taste test held on Youtube. Viewers who participated were asked to fill out a survey based on 4 coffee samples they received. 

# Load Packages and Data
```{r}
knitr::opts_chunk$set(echo = TRUE)
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
library(tidyverse)
library(ggplot2)
library(dplyr)
```

# Data Source Info
* Link to Github: 
 https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-05-14/readme.md

* Survey was voluntary
* Taste test was a follow along live stream. Essentially everyone who participated had the same tasting experience.


# Questions
* Where do they drink coffee?
* How do they brew their coffee?
* What does the consumer generally add to their coffee?
* What is the preferred coffee from the test? 
* Comparison between results! 

# Info at Glance 
```{r}
names(coffee_survey)


```

# Where do they drink coffee?
```{r echo = FALSE}
coffee_data1 <- data.frame(
  Location = c("At home", "At the office", "At a cafe", "On the go", "None of these"),
  Percentage = c(92, 36, 29, 18, 1)
)
ggplot(coffee_data1, aes(x = reorder(Location, -Percentage), y = Percentage, fill = Percentage)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 4.5) +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  scale_fill_gradient(low = "Medium Purple 1", high = "Medium Purple 4") +  
  labs(
    title = "Where do you Brew Coffee?",
    x = NULL, y = "Percent of Respondents"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "none"  
  )



```


# How do they typically brew coffee?
```{r echo = FALSE}

coffee_data2 <- data.frame(
  Method = c("Pour over", "Espresso", "French press", "Other", 
             "Mr. Coffee", "Cold brew", "Pod Machine", 
             "Cometeer", "Instant coffee", "Bean-to-Cup"),
  Percentage = c(63, 41, 20, 18, 18, 14, 9, 5, 4, 2)
)

ggplot(coffee_data2, aes(x = reorder(Method, Percentage), y = Percentage, fill = Percentage)) +
  geom_col(color = "black") +
  scale_fill_gradient(low = "pink", high = "purple") +
  geom_text(aes(label = paste0(Percentage, "%")), hjust = -0.1, size = 4.5) +
  scale_y_continuous(limits = c(0, 70), expand = c(0, 0)) +
  labs(
    title = "How do you Brew Coffee?",
    x = NULL, y = "Percent of Respondents"
  ) +
  coord_flip() +
  theme(legend.position = "none")

```

# Why do they drink coffee?
```{r echo = FALSE}
why_coffee <- data.frame(
  Reason = c("Tastes Good", "Caffeine", "Ritual", "To Poop", "Other"),
  Percentage = c(94, 57, 54, 13, 5)
)

ggplot(why_coffee, aes(x = reorder(Reason, Percentage), y = Percentage, fill = Percentage)) +
  geom_bar(stat = "identity", width = 0.7, color = "black") +
  geom_text(aes(label = paste0(Percentage, "%")), hjust = -0.2) + 
  coord_flip() +
  labs(
    title = "Why Do You Drink Coffee?",
    x = NULL,
    y = "Percent of Respondents"
  ) +
  scale_fill_gradient(low = "Dark Olive Green 4", high = "Dark Olive Green 1") + 
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"), 
    legend.position = "none" )



```


# Coffee Notes / Taste Test 
## Kenyan beans
* Coffee A - light roast 
* Coffee B - medium roast
* Coffee C - dark roast

## Columbian beans
* Coffee D -  light roast



# What is the preference between Coffee A, B, or C?
```{r}
coffee_survey %>% 
  filter(!is.na(prefer_abc)) %>% 
  ggplot(aes(x = prefer_abc, fill = prefer_abc)) + 
  geom_bar() + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  scale_fill_manual(values = c("tan", "tan3", "saddlebrown")) +
  labs(title = "Coffee Preference: A, B, or C?", 
       x = NULL,
       y = "Count") 

```
  
# What is the preference between Coffee A or D?
```{r}
coffee_survey %>% 
  filter(!is.na(prefer_ad)) %>% 
  ggplot(aes(x = prefer_ad, fill = prefer_ad)) + geom_bar() + scale_fill_manual( values = c("tan", "peachpuff3"))+
  labs( title = "Coffee Preference: A or D?",
    x = NULL,
    y = "Count"
  )


```

# What is the overall preference?
```{r}
 coffee_survey%>% 
 filter(!is.na(prefer_overall)) %>%
 ggplot(aes(prefer_overall)) +
 geom_bar(fill= c("tan", "tan3", "saddlebrown", "peachpuff3")) +
 labs(y='Count',
      x = NULL,
      title='Preferred Coffee Between A, B, C or D?')


```

# What does the consumer generally add to their coffee?
```{r}
 



```


# Why do they drink?
```{r echo = FALSE}
why_coffee <- data.frame(
  Reason = c("Tastes Good", "Caffeine", "Ritual", "To Poop", "Other"),
  Percentage = c(94, 57, 54, 13, 5)
)

ggplot(why_coffee, aes(x = reorder(Reason, Percentage), y = Percentage, fill = Percentage)) +
  geom_bar(stat = "identity", width = 0.7, color = "black") +
  geom_text(aes(label = paste0(Percentage, "%")), hjust = -0.2) + 
  coord_flip() +
  labs(
    title = "Why Do You Drink Coffee?",
    x = NULL,
    y = "Percent of Respondents"
  ) +
  scale_fill_gradient(low = "Dark Olive Green 4", high = "Dark Olive Green 1") + 
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"), 
    legend.position = "none" )



```

```{r}




```


*NOTES*
Introduce what the data is(where it comes from)
-Why is it important?
-What features?
*Next, highlight the questions you will be asking.


*Questions*
Where do people brew their coffee? How is it brewed (make relation?)

What is the most popular coffee roast(gender + age)?

#Between Coffee A, Coffee B, and Coffee C which did you prefer?
#Between Coffee A and Coffee D, which did you prefer?
#Favorite Overall Coffee

What are the preferences between popular coffee between gender?

