---
title: "Great American Coffee Taste Test "
author: "Erica Galvan"
date: "2024-11-26"
output:
  slidy_presentation:
    fig_height: 6
    fig_width: 8
  pdf_document: default
 
---
# Introduction
* Coffee expert, "World Champion Barista" and Youtuber 'James Hoffman' partnered with Cometeer, a flash frozen coffee pod delivery service.  
* Live coffee taste test held on Youtube. Viewers who participated were asked to fill out a survey based on 4 coffee samples they received. 

# Load Packages and Data
```{r}
knitr::opts_chunk$set(echo = TRUE)
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
library(tidyverse)
library(ggplot2)
library(dplyr)
```

# Data Source Info
* Link to Github: 
 https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-05-14/readme.md

* Survey was voluntary
* Taste test was a follow along live stream. Essentially everyone who participated had the same tasting experience.

# Info at Glance 
```{r}
names(coffee_survey)


```

# Questions
* Where/Why do they drink coffee?
* How do they brew their coffee?
* What does the consumer generally add to their coffee?
* What is the preferred coffee from the test? 
* Comparison between results! 



# Where do they drink coffee?
```{r echo = FALSE}
coffee_data1 <- data.frame(
  Location = c("At home", "At the office", "At a cafe", "On the go", "None of these"),
  Percentage = c(92, 36, 29, 18, 1)
)
ggplot(coffee_data1, aes(x = reorder(Location, -Percentage), y = Percentage, fill = Percentage)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 4.5) +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  scale_fill_gradient(low = "Medium Purple 1", high = "Medium Purple 4") +  
  labs(
    title = "Where do you Brew Coffee?",
    x = NULL, y = "Percent of Respondents"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "none"  
  )



```


# Why do they drink coffee?
```{r echo = FALSE}
why_coffee <- data.frame(
  Reason = c("Tastes Good", "Caffeine", "Ritual", "To Poop", "Other"),
  Percentage = c(94, 57, 54, 13, 5)
)

ggplot(why_coffee, aes(x = reorder(Reason, -Percentage), y = Percentage, fill = Percentage)) +
  geom_bar(stat = "identity", width = 0.7, color = "black") +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5) + 
  labs(
    title = "Why Do You Drink Coffee?",
    x = NULL,
    y = "Percent of Respondents"
  ) +
  scale_fill_gradient(low = "Dark Olive Green 4", high = "Dark Olive Green 1") + 
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )

```

# How do they typically brew coffee?
```{r echo = FALSE}

coffee_data2 <- data.frame(
  Method = c("Pour over", "Espresso", "French press", "Other", 
             "Mr. Coffee", "Cold brew", "Pod Machine", 
             "Cometeer", "Instant coffee", "Bean-to-Cup"),
  Percentage = c(63, 41, 20, 18, 18, 14, 9, 5, 4, 2)
)

ggplot(coffee_data2, aes(x = reorder(Method, Percentage), y = Percentage, fill = Percentage)) +
  geom_col(color = "black") +
  scale_fill_gradient(low = "pink", high = "purple") +
  geom_text(aes(label = paste0(Percentage, "%")), hjust = -0.1, size = 4.5) +
  scale_y_continuous(limits = c(0, 70), expand = c(0, 0)) +
  labs(
    title = "How do you Brew Coffee?",
    x = NULL, y = "Percent of Respondents"
  ) +
  coord_flip() +
  theme(legend.position = "none")

```


# What does the consumer generally add to their coffee?
```{r echo = FALSE}
 
coffee_adds <- data.frame(
  CoffeeAdds = c("No - just black", "Milk, Dairy, or Creamer", "Sugar or sweetener", "Flavor syrup", "Other"),
  Percentage = c(66, 43, 13, 6, 1)
)

# Plot
ggplot(coffee_adds, aes(x = Percentage, y = reorder(CoffeeAdds, Percentage), fill = Percentage)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  scale_fill_gradient(low = "Coral 1", high = "Coral 3") + 
  labs(
    title = "What do You Add to Coffee?",
    x = "Percent of Respondents",
    y = ""
  ) +
  theme_minimal() 

```


# Coffee Notes / Taste Test 
## Kenyan beans
* Coffee A - light roast 
* Coffee B - medium roast
* Coffee C - dark roast

## Columbian beans
* Coffee D -  light roast

##Limitations 
* Coffee rating is on a 1 to 5 scale. 
* Ranking the preference of bitterness, acidity and preference.
* No Coffee add ons are allowed on the test. 8oz of boiled water only.



# What is the preference between Coffee A, B, or C?
```{r}
coffee_survey %>%
  filter(!is.na(prefer_abc)) %>%
  ggplot(aes(x = prefer_abc, fill = prefer_abc)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) + 
  scale_fill_manual(values = c("tan", "tan3", "saddlebrown")) +
  labs(
    title = "Coffee Preference: A, B, or C?", 
    x = NULL,
    y = "Count"
  ) +
  theme_minimal()

```
  
# What is the preference between Coffee A or D?
```{r}
coffee_survey %>% 
  filter(!is.na(prefer_ad)) %>% 
  ggplot(aes(x = prefer_ad, fill = prefer_ad)) + 
  geom_bar() + 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) + 
  scale_fill_manual( values = c("tan", "peachpuff3"))+
  labs( title = "Coffee Preference: A or D?",
    x = NULL,
    y = "Count"
  )


```

# What is the most popular coffee?
```{r}
 coffee_survey%>% 
 filter(!is.na(prefer_overall)) %>%
 ggplot(aes(prefer_overall)) +
 geom_bar(fill= c("tan", "tan3", "saddlebrown", "peachpuff3")) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) + 
 labs(y='Count',
      x = NULL,
      title='Preferred Coffee Between A, B, C or D?')


```
# Summary 
* 


