---
title: "Coffee"
author: "Erica Galvan"
date: "2024-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
library(tidyverse)
library(ggplot2)
library(dplyr)
```



```{r}
#Defining each coffee A, B, C, and D by their defined roast level. (Light roast, medium, dark, light
roast_colors = c(
  "peru", #COFFEE A = light roast
  "tan3", #COFFEE B = medium roast
  "tan4", #COFFEE C = dark roast
  "tan" #COFFEE D = light roast 
)

```


```{r}

#QUESTION: WHAT IS THE PREFERRED ROAST LEVEL?"

coffee_survey %>% 
  filter(!is.na(prefer_abc)) %>% 
  ggplot(aes(x = prefer_abc, fill = prefer_abc)) + geom_bar() + scale_fill_manual( values = c("tan", "tan3", "saddlebrown"))
```
  

```{r}
coffee_survey %>% 
  filter(!is.na(prefer_ad)) %>% 
  ggplot(aes(x = prefer_ad, fill = prefer_ad)) + geom_bar() + scale_fill_manual( values = c("tan", "peachpuff3"))
  labs( title = "Coffee Preference: A or D?",
    x = "Preference",
    y = "Count"
  )


```


```{r}




```

*NOTES*
Introduce what the data is(where it comes from)
-Why is it important?
-What features?
*Next, highlight the questions you will be asking.


*Questions*
Where do people brew their coffee? How is it brewed (make relation?)

What is the most popular coffee roast(gender + age)?

#Between Coffee A, Coffee B, and Coffee C which did you prefer?
#Between Coffee A and Coffee D, which did you prefer?
#Favorite Overall Coffee

What are the preferences between popular coffee between gender?

