---
title: "Coffee"
author: "Erica Galvan"
date: "2024-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
library(tidyverse)
library(ggplot2)
```

```{r}

#QUESTION: WHAT IS THE PREFERRED ROAST LEVEL?"
coffee_survey %>% 
  filter(!is.na(prefer_abc)) %>% 
  ggplot(aes(x = prefer_abc, fill = prefer_abc)) + geom_bar() + scale_fill_manual( values = c("brown", "black", "yellow"))
```
  


*Questions*
How can I clean up this dataset to answer my questions in more simple way?




*NOTES*
Introduce what the data is(where it comes from)
-Why is it important?
-What features?
*Next, highlight the questions you will be asking.


*Questions*
Where do people brew their coffee? How is it brewed (make relation?)

What is the most popular coffee roast(gender + age)?

#Between Coffee A, Coffee B, and Coffee C which did you prefer?
#Between Coffee A and Coffee D, which did you prefer?
#Favorite Overall Coffee

What are the preferences between popular coffee between gender?

